\hspace{-0.1cm}{J-ReCoVer}:~Java~Reducer~Commutativity~Verifier

### 2019-06-25 draft.tex
sec 5
Modifications in 4.1

In order to explain the problem with further details,
consider the example in Figure~\ref{fig:reducer_max1}. The reducer presented in
the figure remembers the first input value in the variable $m$, increases its
value by 10, and then updates its value to bigger ones if any occur in the loop.
The main loop of the reducer is commutative in this case, but the reducer is not
commutative. A counterexample can be easily found. With the list $[1,2,3,4,5]$
and its permutation $[5,4,3,2,1]$ as the inputs, the reducer outputs $11$ and
$15$, respectively.

@@

To deal with the problem, we apply a program transformation on the
example from Figure~\ref{fig:reducer_max1}. In particular, we move the prefix
$s_1$ into the loop body and use a new variable $s$ to force that the execution
of $s_1$ is always before the original loop body. The result after the
transformation is demonstrated in Figure~\ref{fig:reducer_max2}. The new reducer
program has the same inputs/outputs as the original one. Therefore, if the new reducer is commutative, the original one~is~also~commutative.

@@

In general, the problem with the multiple inputs before the loop can be handled as follows. Given a reducer $c_0;x_{1}{:=} \cur;c_1; \ldots; x_{m}{:=}\cur;c_m;\rloop\{s_2\};s_3$, where $m$ inputs are read before the loop and $\cur$ does
not occur in $c_0,c_1,\ldots, c_m$. 
In the transformed reducer, the part before the loop is now $c_0;s:=1$. The loop body contains several branch statements. 
For all $j \in [1,m]$, we add a branch statement $\IF{s=j}{x_{j}:=\cur; c_j; s:=s+1}$. 
We also add $\IF{s=m+1}{s_2}$ for the original loop body.

>>>>

To illustrate the issue, we consider the reducer shown in
Figure~\ref{fig:reducer_max1}.
%
The reducer presented in the figure remembers the first input value in the
variable $m$, increases its value by 10, and then updates its value to bigger
ones if any occur in the loop.  The main loop of the reducer is commutative in
this case, but the reducer is not commutative. A counterexample can be easily
found. With the list $[1,2,3,4,5]$ and its permutation $[5,4,3,2,1]$ as the
inputs, the reducer outputs $11$ and $15$, respectively.

@@

Our transformation will handle the example from Figure~\ref{fig:reducer_max1} as
follows.
%
We move the prefix $s_1$ into the loop body and use a new variable $s$ to force
that the execution of $s_1$ is always before the original loop body. The result
after the transformation is demonstrated in Figure~\ref{fig:reducer_max2}. The
new reducer program has the same inputs/outputs as the original one. Therefore,
if the new reducer is commutative, the original one~is~also~commutative.

@@

% In general, the problem with the multiple inputs before the loop can be handled
% as follows. Given a reducer $c_0;x_{1}{:=} \cur;c_1; \ldots;
% x_{m}{:=}\cur;c_m;\rloop\{s_2\};s_3$, where $m$ inputs are read before the loop
% and $\cur$ does not occur in $c_0,c_1,\ldots, c_m$.  In the transformed reducer,
% the part before the loop is now $c_0;s:=1$. The loop body contains several
% branch statements. For all $j \in [1,m]$, we add a branch statement
% $\IF{s=j}{x_{j}:=\cur; c_j; s:=s+1}$.  We also add $\IF{s=m+1}{s_2}$ for the
% original loop body.

In general, the problem with the input before the loop can be handled as
follows, including the case where $\cur$ occurs multiple times in $s_1$. Assume
that the $s_1$ part of the reducer $s_1;\rloop\{s_2\};s_3$ has the form
$c_0;x_{1}{:=} \cur;c_1; \ldots; x_{m}{:=}\cur;c_m$ where $\cur$ does not occur
in $c_0,c_1,\ldots, c_m$. In the transformed reducer, the part before the loop
will be $c_0;s:=1$, and the loop body will contain several new branch
statements. In particular, for all $j \in [1,m]$, we add the branch statement
$\IF{s=j}{x_{j}:=\cur; c_j; s:=s+1}$. Moreover, we transform the original loop
body into the branch $\IF{s=m+1}{s_2}$.
